\documentclass[11pt]{article}

%Don't change any thing before \begin{document}
%In fact if you use sth fancy, you might need
%to add more packages, or macros.
\usepackage{amssymb,amsmath}
\usepackage{times,psfrag,epsf,epsfig,graphics,graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}


\begin{document}
\date{}
%\date{Mar 11, 2020}

\title{CSCI 338: Assignment~4~(7 points)}

%\author{Your Name Here}

%\author{ Name1 \and Name2}
%if you have a collaborator 


\maketitle

%When writing up your solution, comment out the following until you reach Problem 1.
\noindent
This assignment is due on {\bf Thursday, April 1, 8:00pm}. It is strongly
encouraged that you use Latex to generate a single pdf file and upload it
under {\em Assignment 4} on D2L. But there will NOT be a penalty for not
using Latex (to finish the assignment). This {\bf could} be a
group-assignment, and you can form a group of at most 2 people (you must put
both names on the group assignment and it is enough for one of the 2 students
to submit it); certainly, you can also finish the assignment all by yourself.

\section*{Problem 1}

Let ${\cal B}$ be the set of all infinite sequences over $\{a,b\}$. Show that
${\cal B}$ is uncountable, using a proof by diagonalization.


\section*{Problem 2}

Let $T=\{(i,j,k)|i,j,k\in{\cal N}\}$. Show that $T$ is countable.

\section*{Problem 3}

Let $INFINITE_{PDA}=\{<M>|M$ is a PDA and $L(M)$ is an infinite language$\}$.
Show that $INFINITE_{PDA}$ is decidable.
\newline

/* ------------ the above contents have been covered by March 12 ------------ */

\section*{Problem 4}

Let $\Sigma=\{a,b\}$. Define the following language {\em ODD}$_{TM}$:

{\em ODD}$_{TM}=\{ <M>|M$ is a TM and $L(M)$ contains only strings of odd length $\}$.

Prove that {\em ODD}$_{TM}$ is undecidable.

\section*{Problem 5}

Show that $EQ_{CFG}$ is undecidable.
\newline

\section*{Problem 6}

Show that $EQ_{CFG}$ is co-Turing-recognizable.
\newline

\section*{Problem 7}

Problem 5.3 (page 239---third edition of Sipser).
\newline

\end{document}
